Machine learning enhanced empirical potentials for metals and alloys

### 2020-05-26 doc/papers/ml-eam/manuscript.tex
Revision 1 for cpc.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Abstract
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{abstract}
Interatomic potential or force-field plays a vital role in simulating materials. 
Empirical potentials like the embedded atom method (EAM) and its variant 
angular-dependent potential (ADP) have proven successful in modeling many 
metals qualitatively. In the past few years, machine learning has became the 
key approach for modeling interatomic interactions. Powered by big data and 
efficient optimizers, machine learning interatomic potentials can generally 
approximate to the first-principle accuracy. In this works, we successfully 
developed a route to express EAM and ADP within machine learning framework in 
highly-vectorizable form and further enhanced the potentials with physical 
constraints. Machine learning can significantly boost their performances with 
few training data. 
For energy and force 
predictions, machine lear̦ned EAM and ADP can be almost as accurate as the 
spectral neighbor analysis potential (SNAP) on the fcc Ni, bcc Mo and Mo-Ni 
alloy systems, while the molecular dynamics simulations performed by our 
potentials are nearly 1000 times faster. Machine learned EAM and ADP can also 
give first-principle accuracy key material properties, such as elastic 
constants, melting temperatures and surface energies. Our results suggest a new 
and systematic route for developing machine learning interatomic potentials. All 
the new algorithms have been implemented in our program TensorAlloy.
\end{abstract}

@@

For the first step, the implementation is almost the same with our previous 
work \cite{TensorAlloy} except the formula of the potential energy. The direct
computation graph from atomic positions to total energy can be built because of
our virtual atom approach (VAP) \cite{TensorAlloy} and the vectorized EAM theory
\ref{sec:transformation}. Hence, atomic forces and virial stress tensors can be 
calculated automatically thanks to the AutoGrad featue of TensorFlow. The 
batching scheme is adopted in this step.

For the second step, only very few specific 
structures are needed to form physical constraints. Thus, the batching scheme is 
no longer a pre-requiste. Instead, we use the codes of the 'prediction' phase of 
VAP for flexibility \cite{TensorAlloy}. For each type of physical constraint of 
each selected crystal structure, an independent computation graph shall be 
generated by reusing the model parameters. This can be easily achieved within 
our VAP framework. Hence, the complicated derivatives of the losses with respect 
to the model parameters can be solved by TensorFlow automatically.
For example, for the Ni-Mo system, if we select the elastic constants of 
$c_{11}$, $c_{22}$ and $c_{44}$ of fcc Ni and bcc Mo, and the Rose EOS of fcc 
Ni and bcc Mo as the physical constraints, there should be $2 \times 2$ separate 
computation graphs (\textrm{Elastic/fcc/Ni, Elastic/bcc/Mo, Rose/fcc/Ni, 
Rose/bcc/Mo}) under the scope 'Constraint'.

The total loss is just the sum of the two steps introduced above:
Equation \ref{eq:loss} demonstrates the loss function used in 
this work:
\begin{align}
\label{eq:loss}
\mathbf{Loss} & = \sqrt{\frac{1}{N_{b}}\sum_{i=1}^{N_{b}}{\left(
    E_{i} - E_{i}^{\mathrm{dft}}
\right)^2}} \nonumber \\
& + \chi_{\mathrm{f}}\sqrt{
    \frac{1}{3\sum_{i}^{N_{b}}{N_i}}\sum_{i}^{N_b}{\sum_{j}^{N_i}{
        \sum_{\alpha}{
            \left(f_{ij\alpha} - f_{ij\alpha}^{\mathrm{dft}}\right)^2
        }
    }}
} \nonumber \\
& + \chi_{\mathrm{rose}}\mathbf{L}^{\mathrm{Rose}} 
+ \chi_{\mathrm{elastic}}\mathbf{L}^{\mathrm{elastic}}
\end{align} 
where $N_{b}$ is the batch size and $N_i$ is the number of atoms in structure 
$i$. $\chi_{\mathrm{f}}$, $\chi_{\mathrm{rose}}$ and $\chi_{\mathrm{elastic}}$ 
are overall weights of force, rose EOS and elastic contributions.

To find optimal parameters of the funcitons in section \ref{sec:functions}, the 
overall loss function shall be defined. Just like other machine learning tasks, 
the mini-batch stochastic gradient descent algorithm is used to minimize the 
loss function. 
The Adam optimizer \cite{adam} is used to minimize Equation \ref{eq:loss}. 
In most cases, we use 0.01 as the initial learning rate and the batch size 
ranges from 20 to 50. Since there are very few adjustable parameters compared 
with neural networks, the optimization typically needs very few epochs to 
converge. When the optimization is finished, the corresponding Lammps setfl 
potential file will be exported.

@@

To conclude, we have successfully combined empirical potentials (EAM/ADP) with 
machine learning. The machine learning approaches (big data, optimization, etc), 
together with physical constraints (Rose EOS, elastic constants) can 
significantly improve the performaces of EAM/ADP. For the fcc Ni, bcc Mo and 
Mo-Ni alloys, ML-EAM and ML-ADP can be as accurate as SNAP, while later method 
is orders of magnitude slower. Our work also indicates a new route to design and 
develop machine learning interatomic potentials \textemdash the machine learning 
enhanced empirical potential approach.

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Author Contributions
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\section*{Author Contributions}
X.C. designed the research project, wrote the related codes and build the 
machine learning interatomic potentials. L.F.W did the molecular dynamics 
simulations and calculations. X.Y.G., Y.F.Z., D.Y.L., W.D.C. and H.F.S. provided 
valuable comments and suggestions to the work. X.C. and L.F.W. wrote the paper. 
All authors contributed to discussions of the results.

>>>>

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Abstract
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{abstract}
Interatomic potential (i.e. force-field) plays a vital role in atomistic 
simulation of materials. 
Empirical potentials like the embedded atom method (EAM) and its variant 
angular-dependent potential (ADP) have proven successful in modeling many 
metals qualitatively. In the past few years, machine learning has became a 
compelling approach for modeling interatomic interactions. Powered by big data and 
efficient optimizers, machine learning interatomic potentials can generally 
approximate to the the accuracy of the first-principles calculations based on 
the quantum mechanics theory. In this works, we successfully 
developed a route to express EAM and ADP within machine learning framework in 
highly-vectorizable form and further incorperated several physical constraints 
into the training. As it is proved in this work, the performances of empirical 
potentials can be significantly boosted with few traning data.
For energy and force 
predictions, machine lear̦ned EAM and ADP can be almost as accurate as the 
spectral neighbor analysis potential (SNAP) on the fcc Ni, bcc Mo and Mo-Ni 
alloy systems, while the molecular dynamics simulations performed by our 
potentials are nearly 1000 times faster. Machine learned EAM and ADP can also 
reproduce some key materials properties, such as elastic constants, melting 
temperatures and surface energies, close to the first-principles accuracy. 
Our results suggest a new and systematic route for developing machine learning 
interatomic potentials. All the new algorithms have been implemented in our 
program TensorAlloy.
\end{abstract}

@@

For the first step, the implementation is almost the same with our previous 
work \cite{TensorAlloy} except the formula of the potential energy. Briefly, 
this step starts from checking through the training dataset and determine some
key constants, including $N^{vap}$ and $N^{\mathrm{max}}_{\mathrm{ij}}$. 
$N^{\mathrm{max}}_{\mathrm{ijk}}$ is not needed as neither EAM nor ADP contains 
triple-atoms interaction. Then we can build the global symbol list (GSL). Hence,
structures of any stoichiometries can be transformed to their corresponding 
unversal GSL forms for batching. The descriptors (\ref{sec:transformation}) are 
built upon these GSL arrays. Then the direct computation graph from atomic 
positions (GSL) to total energy can be easily built. Hence, atomic forces and 
virial stress tensors can be calculated automatically thanks to the AutoGrad 
feature of TensorFlow.

For the second step, the situation is quite different. Only a few structures, 
which may vary significantly, will be used to form the physical constraints. 
Hence, batching is no longer a concern for this step. Instead, we need a more 
flexible way to model complicated physical equations. 
Recall that the virtual-atom approach has two phases: training and prediction 
\cite{TensorAlloy}. The training phase is designed for mini-batch training. 
$N^{vap}$, $N^{\mathrm{max}}_{\mathrm{ij}}$ and the GSL are pre-determined 
constants. On the other hand, the prediction phase targets on a single input 
structure. $N^{vap}$, $N^{\mathrm{max}}_{\mathrm{ij}}$ and the GSL are 
determined dynamically. Therefore, the prediction phase is much suitable for 
modeling physical constraints. In our current implementation, for simplicity, we 
just generate an independent compuation graph for each type of physical 
constraint of each selected crystal structure. The model parameters are shared 
across all the computation graphs (including the graph of step 1). For 
example, for the Ni-Mo system, if we select the elastic constants of $c_{11}$, 
$c_{22}$ and $c_{44}$ of fcc Ni and bcc Mo, and the Rose EOS of fcc Ni and bcc 
Mo as the physical constraints, there should be $2 \times 2$ separate 
computation graphs (\textrm{Elastic/fcc/Ni, Elastic/bcc/Mo, Rose/fcc/Ni, 
Rose/bcc/Mo}) under the scope 'Constraint'. The complicated derivatives of the 
losses with respect to the model parameters can be solved by the AutoGrad module 
of TensorFlow automatically.

The total loss is just the sum of the two steps introduced above:
Equation \ref{eq:loss} demonstrates the loss function used in 
this work:
\begin{align}
\label{eq:loss}
\mathbf{Loss} & = \sqrt{\frac{1}{N_{b}}\sum_{i=1}^{N_{b}}{\left(
    E_{i} - E_{i}^{\mathrm{dft}}
\right)^2}} \nonumber \\
& + \chi_{\mathrm{f}}\sqrt{
    \frac{1}{3\sum_{i}^{N_{b}}{N_i}}\sum_{i}^{N_b}{\sum_{j}^{N_i}{
        \sum_{\alpha}{
            \left(f_{ij\alpha} - f_{ij\alpha}^{\mathrm{dft}}\right)^2
        }
    }}
} \nonumber \\
& + \chi_{\mathrm{rose}}\mathbf{L}^{\mathrm{Rose}} 
+ \chi_{\mathrm{elastic}}\mathbf{L}^{\mathrm{elastic}}
\end{align} 
where $N_{b}$ is the batch size and $N_i$ is the number of atoms in structure 
$i$. $\chi_{\mathrm{f}}$, $\chi_{\mathrm{rose}}$ and $\chi_{\mathrm{elastic}}$ 
are overall weights of force, rose EOS and elastic contributions. Typicall 
$\chi_{\mathrm{f}}$ is 2, $\chi_{\mathrm{rose}}$ is 10 and 
$\chi_{\mathrm{elastic}}$ is set to 0.01.

To find optimal parameters of the funcitons in section \ref{sec:functions}, the 
overall loss function shall be defined. Just like other machine learning tasks, 
the mini-batch stochastic gradient descent algorithm is used to minimize the 
loss function. 
The Adam optimizer \cite{adam} is used to minimize Equation \ref{eq:loss}. 
In most cases, we use 0.01 as the initial learning rate and the batch size 
ranges from 20 to 50. Since there are very few adjustable parameters compared 
with neural network potentials, the optimization typically needs very few 
training steps to converge. When the optimization is finished, the corresponding 
Lammps setfl potential file will be exported.

@@

To conclude, we have successfully combined the training of empirical potentials 
with the machine learning approach into our TensorAlloy framework. The machine 
learning approaches (big data, optimization, etc), 
together with physical constraints (Rose EOS, elastic constants) can 
significantly improve the performaces of EAM/ADP. For the fcc Ni, bcc Mo and 
Mo-Ni alloys, ML-EAM and ML-ADP can be as accurate as the SNAP method, while 
later method is about three orders of magnitude slower. Our work also indicates 
a new route to design and 
develop machine learning interatomic potentials \textemdash the machine learning 
enhanced empirical potential approach.

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Author Contributions
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\section*{Author Contributions}
D.Y.L and H.F.S contributed the central conceptions of this study, designed 
the research project, conducted the analyses and revised the manuscript. X.C. 
designed the research project, wrote the related codes, trained the machine 
learning interatomic potentials and conducted the analyses. L.F.W performed 
molecular dynamics simulations and validated our models. X.Y.G., Y.F.Z. and 
W.D.C: provided valuable comments and suggestions to the work. X.C. and L.F.W. 
wrote the initial draft of the paper. All authors contributed to discussions of 
the results.
